### 2019-03-01 results.tex
mz - Michael's pass

\begin{table}[t]
\centering
\input{tables/application-tab}
\caption{Machine learning applications used as benchmarks
to evaluate Seneca. 
%paper submission is blind (so such references must be omitted so as not to reveal our identities):
%Code base is available at project repository~\cite{ref:seneca}.
\label{tab:bmarks}}
%\vspace{-0.2in}
\end{table}

\begin{table}[t]
\centering
\input{tables/hparams-prophet-tab}
\caption{The hyperparameters that Seneca considers for Prophet. 
The default value and tuning options are listed. 
\label{tab:prophet_para}}
%\vspace{-0.2in}
\end{table}

\begin{table}[t]
\centering
%\scriptsize
\input{tables/hparams-xgboost-tab}
\caption{The hyperparameters that Seneca considers for XGBoost. 
The default value and tuning options are listed. 
\label{tab:xgboost_para}}
%\vspace{-0.2in}
\end{table}

\begin{table}[t]
\centering
%\scriptsize
\input{tables/hparams-svc-tab}
\caption{The hyperparameters that Seneca considers for SVC. 
The default value and tuning options are listed. 
\label{tab:svc_para}}
%\vspace{-0.2in}
\end{table}

\begin{table}[t]
\centering
%\scriptsize
\input{tables/hparams-nn-tab}
\caption{The hyperparameters that Seneca considers for NN.
The default value and tuning options are listed. 
\label{tab:nn_para}}
\vspace{-0.2in}
\end{table}

In this section,
we evaluate Seneca in terms of performance, monetary cost, 
and hyperparameter tuning execution time, 
for different machine learning applications.
We first overview these applications and our empirical methodology. 
We then present our experimental results.

\subsection{Benchmark Applications and Training/Testing  Datasets}
The machine learning applications that we use to evaluate Seneca 
are shown in Table~\ref{tab:bmarks}, each with 
a brief description.  Prophet, Multi-Regression, and XGBoost are regression
applications; XGBoost, SVC, and NN are classification applications (XGBoost 
implements both regression and classification tasks).

Prophet~\cite{ref:prophet} is a time series analysis library built and maintained by Facebook and the open-source community. The input dataset is a time series of view counts
%logorithm
of Peyton Manning's Wikipedia page from Dec. 10th, 2007 to Jan. 20th, 2016. 
The dataset exhibits both seasonality and a holiday effect (e.g. around 
super bowl games). 
We use the first 6 years as the training set and the last 2 years as the testing set.  We use a cross-validation horizon (sliding window) of 1-year, 
and a period (sliding pace) 
of 180 days.  As such, Seneca performs three cross-validations for each 2-year test range.

Prophet expects multiple hyperparameters: \textit{growth} specifies linear or logistic trend model growth; \textit{prior scale} indicates the strength of the 
sparse prior probability~\cite{ref:sparse_prior}. 
There are three prior scale hyperparameters for change point, holidays, and seasonality. 
Since Prophet uses a Fourier sum to estimate seasonality, 
the \textit{fourier order} is the number of terms in the partial Fourier 
sum. \textit{Seasonality mode} indicates that the effect of seasonality is either 
multiplicative or additive. Finally, the width of uncertainty intervals 
is set using \textit{interval width}.

In the results that follow, we evaluate \textit{default} parameterizations (that come with 
the application or that are defined by the application developer) for each application.  
The default parameters and tuning options for Prophet are listed in the Table~\ref{tab:prophet_para}. For every cross validation, the application computes mean square error (MSE) as $\frac{1}{n}\sum_{i=1}^{n}(Y_i - \hat{Y_i})$, where $\hat{Y_i}$ is the ground truth, $Y_i$ is model prediction and n is the number of data points. Then the application returns the average MSE across cross validations. 
%Based on the model trained by default parameter setting, this average MSE metric is 0.284.

Multi-Regression is a regression application developed as part of
an Internet-of-Things (IoT) project~\cite{iot-cpu} (which has been
extended from linear regression described in the citation 
to multiple linear regression by the authors of this prior work).
The application predicts outdoor temperature from the processor 
temperature of single board computers (SBCs).  
The technique uses multiple linear regression models computed from time
series of processor temperature measurements,
to predict outdoor microclimate temperatures. 
The training dataset consists of eight input time 
series (one per SBC, each containing 
5-minute measurements) from April 5th, 2018 to Dec. 10th, 2018.

A hyperparameter configuration for this application is 
a subset of input SBC time series.
Seneca considers all \texttt{$2^N - 1$} potential subsets (for $N$ input time series).
For this application,
we assume the default parameterization is the 
full set of input time series (8 in this case).
The test dataset for this application is a time 
series of the outdoor temperature (ground truth) 
over the same period.  The application makes 
predictions for each of these outdoor temperatures
using the regression coefficients constructed from the training set
and computes and returns the MSE between predicted 
and ground truth value pairs.

XGBoost~\cite{ref:xgboost-web} is an open source 
framework for gradient boosting, which 
performs both regression and classification. 
The hyperparameters and the default values that Seneca uses 
in this study are listed in Table~\ref{tab:xgboost_para}. 
A definition for each hyperparameter can be found in~\cite{ref:xgboostparams}.
%xgboostparams: https://xgboost.readthedocs.io/en/latest/python/python_api.html#module-xgboost.sklearn

Support vector classification (SVC)~\cite{ref:svc} 
is an implementation of a classification algorithm 
that is based on support vector machines that is part of the
libsvm~\cite{ref:libsvm} library.
The hyperparameters and their defaults that Seneca uses for SVC in this study are listed in Table~\ref{tab:svc_para}. A definition for each hyperparameter can be found in
%svcparams: https://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html
\footnote{\url{https://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html}}.

Neural network (NN)~\cite{ref:neural_network} is a machine 
learning framework that learns 
and makes decisions from patterns in an input dataset. For this application, 
we use a feed-forward multi-layer perceptron model~\cite{ref:feedforward_nn} 
for the classification task.
The hyperparameters and their defaults for NN 
are listed in Table~\ref{tab:nn_para}. 
A definition for each hyperparameter can be found in~\cite{ref:nnparams}.
%nn: https://scikit-learn.org/stable/modules/generated/sklearn.neural_network.MLPClassifier.html

For the classification applications (XGBoost, SVC, and NN), we use a labeled
dataset for training, testing, and evaluation from an IoT 
project~\cite{blind}. The dataset contains measurements from individual
citrus fruit (e.g. oranges, mandarins, lemons, etc.) taken by a fruit sorting
and grading device (using a large number of sensors).  The measurements (i.e.
features) include size, shape, weight, color, diameter, flatness, among other
characteristics, for each fruit.  
The dataset has been filtered to remove correlated
features (those with an absolute value of the Pearson correlation coefficient
greater than 0.8). The dataset has also been balanced by 
down-sampling and the resulting 
dataset contains 33926 rows (individual fruit) 
distributed evenly across 5 targets. Each row has
18 features.  The label identifies the field from which the 
individual fruit was harvested.

The applications train a model on a subset (80\%) of the
data.  Each then uses this model to predict the field
from which each fruit originates (for the remaining 20\%). 
The applications compute a classification accuracy score
as $\frac{1}{n}\sum_{i=1}^{n}1(Y_i = \hat{Y_i})$, where $Y_i$ is the
prediction class, $Y_i$ is the true class, 
$n$ is the number of samples, and 1(x)
is the indicator function. The applications return the accuracy percentage
upon completion. As mentioned above, in addition to returning
classification accuracy for its classification task, XGBoost also returns 
MSE for its regression task.

%The dataset is from a Compac packline machine~\cite{ref:compac} that sorts and grades fruit using a large number of 
%measurements (i.e. feature measurements including size, shape, weight, color, diameter, flatness, etc.).  The labels identify the field from which fruit was harvested.
%The dataset is filtered to remove correlated features (those with an absolute value of the Pearson correlation 
%coefficient greater than 0.8).

%Seneca downloads the dataset from AWS S3 for each function invocation.
%It then randomly indexes the dataset and splits it into two.

% For these applications (XGBoost, SVC, and NN), we consider three
% split scenarios for training and testing: 80\%-20\%, 20\%-80\% and
% 0.1\%-99.9\%. As the proportion of training data decreases, the performance of
% classifier degrades accordingly. However we note that using the optimized
% hyperparameters generated by Seneca, these classifiers are able to achieve
% accuracy levels similar to those generated from the default parameters, but
% with significantly less training data.

% and we argue that Seneca is able to eke out
%improvement even with very limited training data.

%\textcolor{blue}{list them here once we finish the additional results} 
%This setup also means that each Seneca function
%considers a different training and testing set during hyperparameter tuning.

%todo in the future -- for a complete tuning, use the same testing and training sets
%run this multiple times (for the same test/train datasets) and report the average and stdev in the latency (and memory?) results.
%todo in the future -- repeat the above using different test/train datasets, but keep 
%them the same for all functions in a single tuning run.  Compare the MSE/accuracy and
%the configurations that are selected, analyze how they change (if they do)

\subsection{Empirical Methodology}

To evaluate Seneca, we measure tuning performance, application execution time, 
memory use, and monetary cost.  For tuning performance, 
we use mean squared error (MSE)
for regression tasks and percentage accuracy for classification 
tasks (i.e. we score model prediction accuracy and not explanatory power).  
Seneca is able to compare models for each type of application using 
these scores because it uses the same
total number of hyperparameters for each model.  Thus, the
the penalty function in terms of 
Bayesian Information Criteria is the same.

The default configuration that we consider for each application is described
above.  We report results for the default, best (Seneca's recommendation), and
worst performing hyperparameter configurations for each application type.  
Seneca computes all possible combinations of
the hyperparameter settings specified in the configuration to extract each of
these results.  The default results represents those that a novice or first
time user might employ when using these applications as a ``black box.''  The
worst performing scenario shows how bad the results can be when parameters are
poorly tuned.  Finally, the best is the upper bound on what is possible from
tuning the hyperparameters for the values and dataset specified 
(e.g. using expert knowledge or Seneca).

To evaluate the  savings introduced by Seneca, we deploy the applications
on Amazon Web Services (AWS) Lambda and extract
execution time and memory use from AWS CloudWatch~\cite{ref:awscloudwatch} logs.  
Seneca extracts execution time and memory use from AWS CloudWatch logs\footnote{\url{https://aws.amazon.com/cloudwatch/}}.
We compute monetary cost
using the AWS Lambda pricing model\footnote{\url{https://aws.amazon.com/lambda/pricing/}}.
%awslambdapricing: https://aws.amazon.com/lambda/pricing/
Each function downloads the training and testing dataset 
of the application from AWS S3 upon function invocation. 
We do not consider the cost of dataset storage 
in our cost computations (it is very small).

%We also experiment with different memory configurations for Seneca functions
%in AWS Lambda.  
We also investigate Seneca's automatic memory optimization capabilities.  To
do so we compare the execution performance and cost of the benchmarks using
the maximum allocatable memory size to the performance and cost when run with
Seneca's automatically determined memory size. Though memory size used reported by AWS CloudWatch fluctuates throughout the tuning process, we 
have verified that the optimized allocated memory is sufficient for all 
hyperparameter configurations to complete successfully.

% Note that this comparison assumes that the memory size used by each benchmark does not depend on hyperparameter settings. We have verified that this assumption holds for the benchmarks and datasets we consider and plan to consider applications for which hyperparameter settings require different maximum amounts of memory as part of future work.

@@

We begin by evaluating Seneca's effect on application efficacy.
That is, we detail the effect on the quality of the output generated by each
application when Seneca determines the hyperparameter settings.

\begin{table}
\centering
\input{tables/mse-tab}
\caption{Combination count and MSE for the default, best (Seneca's recommendation), and worst hyperparameter configurations for the three regression applications. 
For the MSE  values (rows 3-5), lower is better.
\label{tab:mse}}
%\vspace{-0.2in}
\end{table}

\begin{figure}[t] \centering 
\vspace{-0.2in}
\includegraphics[scale=0.36]{box_plot_mse_2}
\vspace{-0.4in}
\caption{The box plot of MSE metric from three regression applications in the entire hyperparameter tuning search space. The red notch shows the MSE from the default settings. The colored diamonds are outliers beyond two interquartile ranges. Lower MSE values are better.
\label{fig:box_plot_mse}}
\vspace{-0.2in}
\end{figure}

%using the three regression applications: 
%Prophet, Multi-Regression, and XGBoost (regression task). 
Table~\ref{tab:mse} shows the effect of Seneca on the quality of the
regressions generated by Prophet, Multi-regression, and XGBoost (when used as
a regression algorithm). The first row indicates the number of
different hyperparameter settings Seneca considered.  Rows 3 through 5 show
the MSE generated by the default hyperparameter settings, the best (lowest
MSE) Seneca was able to find, and the worst (largest MSE) that Seneca tested
for each regression algorithm and dataset.

%s mentioned previously, the table shows regression
%quality in terms of MSE. 
%
%size of the hyperparameter search space (number of configurations
%considered) and the mean squared error (MSE)
%for each of the regression applications that we consider (columns 3-5).
%The first row of data is 
% the number of hyperparameter configurations in the search space of each. 
%The last three rows show the MSE for the default, worst, and best performing 
%hyperparameter configuation reported by Seneca (lower is better).

%Note that Seneca's recommendation is the best performing configuration.
Compared to the default hyperparameter settings,
Seneca reduces MSE by 22.56\%, 15.94\%, and 62.69\%, for Prophet, Multi-Regression, and XGBoost, respectively, for the datasets and training methodologies that we consider.
Verses the worst case, Seneca reduces MSE by 82.62\%, 78.01\%, and 99.51\%, respectively.

Figure~\ref{fig:box_plot_mse} provides a box plot of MSE (lower is better) for
these applications for the entire hyperparameter search space. The central
rectangle covers the interquartile range (IQR), which is defined as the range
of data points from first quartile to third quartile (Q3 - Q1).  The upper
whisker extends to the last datum less than \texttt{$(Q3 + 2 * IQR)$} and the
lower whisker extends to the first datum greater than \texttt{$(Q1 - 2 *
IQR)$}. The data points beyond the whiskers are considered outliers and are
plotted as colored diamonds.

The red notch identifies the MSE that results
from training the model using the default settings of hyperparameter. The
difference between red notch of lowest data point is improvement brought about
through the use of Seneca. From the location of red notch, we can see the MSE
metrics from default configuration in three applications are similar to those
in the first quartile for all applications.
This data shows that the quality of the
output generated by each application (for a given dataset) is dramatically
affected by hyperparameter settings.  Predictably, the default settings are
near the ``good'' end of the spectrum, but as both Table~\ref{tab:mse} and
Figure~\ref{fig:box_plot_mse} indicate, Seneca is able to find
parameterization that improves output quality over the default settings
in each case.
%Thus for Prophet
%and Multi-Regression, Seneca 
%, meaning the potential room for improvement is about 75\%
%over the range of MSE metric. The box plot also illustrates that the best MSE
%of Prophet and multivariate regression are among outliers, where a
%comprehensive search is very critical to find the best configurations. 
%\textcolor{blue}{Add more analysis here to discuss the variance in the MSE
%for the range of configurations and how they change across applications.
%Include any insights that the boxplots add.}

%%%%%%%%%%%%%%%%%%%%% classification apps - tuning performance
%%%%%%%%%%%%%%%%%%%%%%%%%%
We next empirically evaluate the tuning performance of Seneca for the three
classification applications: XGBoost (classification), SVC, and NN.
Table~\ref{tab:accuracy} presents the accuracy percentage (higher is
better) reported by each application (3 rightmost columns) for the default,
worst, and best hyperparameter tuning configurations reported by Seneca.
using an 80/20 (train/test) percent split,
Seneca increases accuracy by 2.46\%, 19.04\%, and 3.79\%, for XGBoost
(classification), SVC, and NN applications, respectively.  Compared to
the worst case, Seneca improves accuracy by 98.11\%, 26.74\%, and 63.92\%,
respectively.

\begin{table}
\centering
\input{tables/accu-tab}
\caption{The Accuracy reported for the 
default, best (Seneca's recommendation), and worst hyperparameter configurations for 
the three classification applications using 80\% of the data to train and 20\%
of the data as a test set. 
%under three train/test split scenarios.  
For the accuracy values in the table, higher is better.
\label{tab:accuracy}}
\vspace{-0.2in}
%\end{figure}
\end{table}

%not sure what this paragraph is adding, seems to a repetition of what
%is already said above...
%The 80-20 case represents the typical usage of these classification
%applications. Often, when evaluating the accuracy of a specific classifier,
%the user uses the 80-20 split ``rule'' when measuring accuracy. In this case,
%Seneca's improvement over the default parameterizations varies from 2.46\% for XGBoost application (column 2), to 19.04\% for SVC
%(column 3).

\begin{figure}[t] \centering 
\includegraphics[scale=0.36]{box_plot_Accuracy_20}
\vspace{-0.4in}
\caption{The box plot of accuracy metric from three classification applications for the entire hyperparameter tuning search space. The red notch indicates the accuracy metric from the trained model using default settings. The colorful diamonds are outliers beyond two interquartile ranges. Higher accuracy values are better.
\label{fig:box_plot_accuracy}}
\vspace{-0.1in}
\end{figure}

Figure~\ref{fig:box_plot_accuracy} presents the accuracy (higher is better) 
for the entire hyperparameter search space as a box plot for each classification
application.  The placement of the rectangle and whiskers is the same
as in Figure~\ref{fig:box_plot_mse}. 
%The central rectangle covers from first quartile to third quartile (Q3 - Q1) and the whiskers span from \texttt{$(Q3 + 2 * IQR)$} to \texttt{$(Q1 - 2 * IQR)$}. 
The red notch indicates the accuracy metric from the model trained by default settings and colored diamonds are outliers beyond whiskers.
Again, Seneca significantly
improves tuning performance over the default and worst case configurations for SVC while generating a smaller improvement for XGBoost and Neural Network versus the default.
This is because XGBoost and NN use a well tuned default parameter
set that works well for most datasets.  SVC benefits more from Seneca tuning
that the others.

\begin{table}[t]
\centering
\scriptsize
\input{tables/exec-time-memory-tab}
\caption{The mean and standard deviation (in parentheses) values of execution time and memory use, and best accuracy across 30 executions of three classification benchmarks running against two random states. It verifies data split does not have major impact to execution metrics and scores.
\label{tab:exec_memory}}
%\vspace{-0.2in}
\end{table}

To investigate the potential impact of the 80-20 data split for the classification
applications, we next evaluate the quality of the output generated from 
each when we consider different 80-20 random splits.
To enable this, we run Seneca 30 times to obtain execution time, memory use, 
and best accuracy score. 
Table~\ref{tab:exec_memory} shows the mean and standard deviation of these 
metrics across runs for two different random splits.  The Seneca score
is similar across splits.  As a result, users can employ the models
(for the selected hyperparameter settings) that Seneca recommends
repeatedly for inference to amortize the cost of using Seneca.

% \begin{table}
% \centering
% \input{tables/speedup-tab}
% \caption{The execution time of tuning in seconds by using Seneca and EC2 instance (Serial) and respective speedup. 
% \label{tab:accuracy}}
% \vspace{-0.2in}
% \end{table}

Finally, we compare Seneca to AWS EC2 use.
We measure the execution time of Seneca and hyperparameter tuning 
using the least expensive EC2 instance type in which the application
will run (t2.micro for SVC and NN, t1.medium for the rest).  End to end, Seneca's
concurrency facilitates in a 4x -- 17x (9x on average) speedup
and \$0.01--\$0.34 (\$0.13) additional cost over serialized execution in EC2.
EC2 trades off monetary cost for programmer productivity, i.e. the developer 
must setup, deploy, and manage the applications manually. Moreover, the EC2
pricing model is imposed on hour boundaries. Moreover, Seneca and 
AWS Lambda impose no cost when no execution is occuring (EC2 instances must
stay idle in between runs). As part of future work, we will compare Seneca
to the AWS Elastic Container Service.
%\textcolor{blue}{FIX THIS: State the results (speed up) per application and 
%also state the cost analysis for 1 hour of computation.  Clarify that this
%is due to the 1-hour pricing model of EC2.}

%We use instance type t2.medium
%for Prophet (since it cannot be built
%process on EC2 instance across five applications. Since library compilation issue, we use t2.medium on Prophet and t2.micro on other 4 applications. 
%As a result, Seneca achieves at least 4x speedup for IO-bound application (SVC) and 21x speedup for CPU-bound application (XGBoost). It confirms Seneca can execute tuning within a more reasonable period of time, comparing to executing serially on EC2 instance.

>>>>

\begin{table}[t]
\centering
\input{tables/application-tab}
\caption{Machine learning applications used 
to evaluate Seneca. 
%paper submission is blind (so such references must be omitted so as not to reveal our identities):
%Code base is available at project repository~\cite{ref:seneca}.
\label{tab:bmarks}}
\vspace{-0.1in}
\end{table}

\begin{table}[t]
\centering
\input{tables/hparams-prophet-tab}
\caption{Hyperparameters Seneca considers for \textbf{Prophet}. 
\label{tab:prophet_para}}
%\vspace{-0.1in}
\end{table}

\begin{table}[t]
\centering
\input{tables/hparams-xgboost-tab}
\caption{Hyperparameters Seneca considers for \textbf{XGBoost}. 
\label{tab:xgboost_para}}
\vspace{-0.2in}
\end{table}

\begin{table}[t]
\centering
%\scriptsize
\input{tables/hparams-svc-tab}
\caption{Hyperparameters Seneca considers for \textbf{SVC}. 
\label{tab:svc_para}}
%\vspace{-0.2in}
\end{table}

\begin{table}[t]
\centering
%\scriptsize
\input{tables/hparams-nn-tab}
\caption{Hyperparameters Seneca considers for \textbf{NN}.
\label{tab:nn_para}}
\vspace{-0.2in}
\end{table}

In this section, we empirically evaluate Seneca in terms of machine learning (ML) 
model output quality, performance, and cost.
We first overview the ML applications 
that we consider and our experimental methodology. 
We then present our results.

\subsection{Benchmark Applications and Training/Testing Datasets}
The ML applications that we use to evaluate Seneca 
are described in Table~\ref{tab:bmarks}.
Prophet, Multi-Regression, and XGBoost are regression
applications; 
XGBoost, SVC, and NN are classification applications (XGBoost 
implements both regression and classification tasks).
The regression applications compute the mean square error (MSE) 
as $\frac{1}{n}\sum_{i=1}^{n}(Y_i - \hat{Y_i})$, where $\hat{Y_i}$ is the ground truth, 
$Y_i$ is model prediction and n is the number of data points. The applications
return the average MSE across cross validations.
The classification applications compute and return a classification accuracy 
percentage, which is calculated
as $\frac{1}{n}\sum_{i=1}^{n}1(Y_i = \hat{Y_i})$, where $Y_i$ is the
prediction class, $\hat{Y_i}$ is the true class, $n$ is the number of samples, and 1(x) is the indicator function.

Prophet~\cite{ref:prophet} is an open source time series analysis library developed
by Facebook.  The input dataset we consider is a time series of view counts
of Peyton Manning's Wikipedia page (Dec. 2007--Jan. 2016).
The dataset exhibits both seasonality and a holiday effect (e.g. around 
super bowl games).  We use the first 6 years as the training set 
and the last 2 years as the testing set.  We use a cross-validation 
horizon (sliding window) of 1-year, and a period (sliding pace) 
of 180 days.  As such, Seneca performs three cross-validations for 2-year test range.

Prophet expects multiple hyperparameters.
\textit{growth} specifies linear or logistic trend model growth and \textit{prior scale} indicates the strength of the sparse prior probability. There are three prior scale hyperparameters for change point, holidays, and seasonality. Since Prophet uses a Fourier sum to estimate seasonality, the \textit{fourier order} is the number of terms in the partial Fourier sum. \textit{Seasonality mode} indicates that the effect of seasonality is either multiplicative or additive. Finally, the width of uncertainty intervals is set using \textit{interval width}.

Each application has default hyperparameter settings (i.e. default values or those recommended by the application maintainer). The hyperparameters, their default and optional values that we consider for Prophet are listed in the Table \ref{tab:prophet_para}.

Multi-Regression is a regression application developed by others as part of
an Internet-of-Things (IoT) project~\cite{iot-cpu} (which has been
extended from linear regression described in the citation 
to multiple linear regression by the authors of this prior work).
The application uses multiple linear regression models
to predict outdoor temperature from the processor 
temperature of single board computers (SBCs).  
The training dataset consists of eight input time 
series (one per SBC, each containing 
5-minute measurements) from Apr. 5th to Dec. 10th, 2018.

Hyperparameter configuration for Multi-Regression is a subset of input SBC time series.
Seneca considers all \texttt{$2^N - 1$} potential subsets (for $N$ input time series).
For this application, the default parameterization is the 
full set of input time series (8 in this case).
The test dataset is a time 
series of the outdoor temperature (ground truth) 
over the same period.  
The application makes 
predictions for each of these outdoor temperatures
using the regression coefficients constructed from the training set
for each new value in the test set.
%and computes and returns the MSE between predicted 
%and ground truth value pairs.

XGBoost~\cite{ref:xgboost-web}, SVC~\cite{ref:svc}, and NN~\cite{ref:neural_network} 
are the classification applications that we consider.
XGBoost~\cite{ref:xgboost-web} is an open source 
framework for gradient boosting, which 
performs both regression and classification. 
The hyperparameters and their default values 
are listed in Table~\ref{tab:xgboost_para}, their definitions can be
found in~\cite{ref:xgboostparams}. SVC uses support vector machines to implement classification as part of the libsvm~\cite{ref:libsvm} library.
The hyperparameters and their defaults that Seneca uses for SVC in this study are listed in Table~\ref{tab:svc_para} with definitions
in~\cite{ref:svcparams}. NN is a machine 
learning application leveraging neural network to identify patterns from an input dataset. 
Here we implement a feed-forward multi-layer perceptron model~\cite{ref:feedforward_nn} 
for classification. The hyperparameters and their defaults for NN are listed in Table~\ref{tab:nn_para} with definitions
in~\cite{ref:nnparams}. 
%\footnote{\url{https://scikit-learn.org/stable/modules/generated/sklearn.neural_network.MLPClassifier.html}}.
%\footnote{\url{https://xgboost.readthedocs.io/en/latest/python/python_api.html\#module-xgboost.sklearn}}
%\footnote{\url{https://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html}}

For these classification applications, we use a labeled
dataset for training, testing, and evaluation from another IoT 
project~\cite{blind}. The dataset contains measurements of individual
citrus fruit (e.g. oranges, mandarins, lemons, etc.) taken by a fruit sorting
and grading device using a large number of sensors. The measurements (i.e. features) include size, shape, weight, color, diameter, flatness, among other characteristics, for each fruit.  
The dataset has been filtered to remove correlated
features (those with an absolute value of the Pearson correlation coefficient
greater than 0.8). The dataset has been balanced by 
down-sampling and the resulting 
dataset contains 33926 rows (individual fruit) 
distributed evenly across 5 targets. Each row has
18 features.  The label identifies the field from which the individual fruit was harvested.

The applications train a model on a random subset (80\%) of the data. Each then uses this model to predict the field from which each fruit originates for the remaining 20\%. To study the impact of random data split, we consider multiple 80\%/20\% splits in our evaluation.
\subsection{Empirical Methodology}

To evaluate Seneca, we measure model output quality, execution time, memory use, and monetary cost.  For output quality (prediction accuracy) our metrics are mean squared error (MSE) for regression and percentage accuracy for classification as described above. We score model prediction accuracy and not explanatory power ($R^2$) to avoid overfitting.

\ignore{Seneca is able to compare models for each type of application using these scores because it uses the same total number of hyperparameters for each model. Thus, the penalty function in terms of Bayesian Information Criteria is the same. 
comments: BIC is based on parameters, not for hyperparameters.}

We compare results for the default, best (Seneca's recommendation), and worst performing hyperparameter configurations for each application type. Seneca computes all possible combinations of the hyperparameter settings specified in the configuration to extract each of these results.  \texttt{default} represents results that a novice or first time user might experience when using these applications as a ``black box.''  The \texttt{worst} shows how bad the results can be when parameters are poorly tuned.  Finally, the \texttt{best} is the upper bound on what is possible from tuning the hyperparameters for the values and datasets specified  (e.g. using expert knowledge or Seneca).

Seneca deploys the applications automatically over AWS Lambda
and extracts execution time and memory use from 
AWS CloudWatch~\cite{ref:awscloudwatch} logs.
We compute monetary cost using the AWS Lambda pricing model~\cite{ref:pricing}.
Each function downloads the training/testing dataset 
of the application from AWS S3 upon function invocation. 
We do not consider the cost of dataset storage 
in our cost computations (it is very small).
%\footnote{\url{https://aws.amazon.com/cloudwatch/}}
%\footnote{\url{https://aws.amazon.com/lambda/pricing/}}
We also evaluate Seneca's automatic memory optimization capabilities.  To
do so we compare the execution performance and cost of the applications using
the maximum allocatable memory size to the performance and cost when run with
Seneca's automatically determined memory size. Even though \texttt{maximum memory used} reported by AWS CloudWatch can fluctuate, we 
have verified that the optimized allocated memory is sufficient for all 
hyperparameter configurations to complete successfully.  We have also verified
that the memory requirements across hyperparemeter settings do not vary 
significantly. We plan to consider applications for which hyperparameter settings require
different maximum memory sizes as part of future work.

@@

\begin{table}
\centering
\input{tables/mse-tab}
\caption{Hyperparameter configuration count and MSE for the default, 
best (Seneca's recommendation), and worst configurations for the three regression applications. 
For the MSE  values (rows 3-5), lower is better.
\label{tab:mse}}
\vspace{-0.2in}
\end{table}

\begin{figure}[t] \centering 
\vspace{-0.5in}
\includegraphics[scale=0.36]{box_plot_mse}
\caption{Box plot of MSE from the three regression applications across the 
hyperparameter tuning search space. The red notch shows the MSE from the default settings. 
The colored diamonds are outliers beyond two interquartile ranges. 
Seneca selects the points indicated by blue triangle.
Lower MSE values are better. 
\label{fig:box_plot_mse}}
\vspace{-0.2in}
\end{figure}

We first evaluate the quality of the output generated by each
ML applications when Seneca determines the hyperparameter settings. We
first show the results for the regression applications in
Table~\ref{tab:mse}.
The first row of data is the number of hyperparameter 
configurations that Seneca considers for each.
The last three rows show the MSE for the default, worst, and best performing (Seneca's recommendation)
hyperparameter configuation (lower is better).
Seneca reduces MSE by 22.56\%, 15.94\%, and 44.88\%, for Prophet, Multi-Regression, and XGBoost,
respectively, for the datasets and training methodologies that we consider.
Verses the worst case, Seneca reduces MSE by 82.62\%, 78.01\%, and 99.28\%, respectively.

Figure~\ref{fig:box_plot_mse} shows the MSE box plot for the hyperparameter search space for these applications (lower is better). 
The central rectangle covers the interquartile range (IQR),
which is defined as the range of data
points from first quartile to third quartile \texttt{$(Q3 - Q1)$}.
The upper whisker extends to the last datum less
than \texttt{$(Q3 + 2 * IQR)$} and the lower whisker extends to the first datum greater
than \texttt{$(Q1 - 2 * IQR)$}. The data points beyond the whiskers are considered outliers
and are plotted as colored diamonds. The red notch identifies the MSE that results from training
the model using the default settings of hyperparameter.
The blue triangle identifies the MSE of Seneca.
The difference between red notch and blue triangle
is the improvement brought about by the use of Seneca, over 
using the default parameter setting.
The plot also shows
that Prophet and Multi-Regression have a significant number of outliers,
indicating that a comprehensive search is critical to finding the best configurations.

%%%%%%%%%%%%%%%%%%%%% classification apps - ML model output quality %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\centering
\input{tables/accu-tab}
\caption{Accuracy for the 
default, best (Seneca's recommendation), and worst hyperparameter configurations for 
the three classification applications using 80\% of the data to train and 20\%
of the data as a test set. Higher accuracy is better.
\label{tab:accuracy}}
\vspace{-0.1in}
\end{table}

\begin{figure}[t] \centering 
\vspace{-0.1in}
\includegraphics[scale=0.36]{box_plot_Accuracy_20}
\vspace{-0.4in}
\caption{Box plot of accuracy reported for three classification applications across 
the hyperparameter search space. The red notch indicates the accuracy that results
from default hyperparameter values.
The diamonds are outliers beyond two interquartile ranges. 
Seneca selects the points indicated by blue triangle. 
Higher accuracy is better. 
\label{fig:box_plot_accuracy}}
\vspace{-0.2in}
\end{figure}

\begin{table}[t]
\centering
\scriptsize
\input{tables/exec-time-memory-tab}
\caption{The mean and standard deviation (in parentheses) for execution time and memory use 
(across 30 runs),
and best accuracy score for the classification applications using two different random splits. 
\label{tab:exec_memory}}
\vspace{-0.1in}
\end{table}

We next empirically evaluate Seneca's model output quality for the three
classification applications: XGBoost (classification), SVC, and NN.
Table~\ref{tab:accuracy} presents the accuracy percentage (higher is better)
for each application (3 right-most columns)
for the default, worst, and best (Seneca's recommendation) hyperparameter tuning
configurations (data rows 2-4).
The first row 
reports the number of configurations that Seneca considers in its search space.
Using a random 80/20 (train/test) percent split,
Seneca increases accuracy by 2.46\%, 19.04\%, and 3.79\%, for XGBoost
(classification), SVC, and NN applications, respectively.  
Because XGBoost and NN use a well-tuned default parameter set that works well for most datasets, Seneca provides only modest improvements.
Compared to the worst case however, Seneca improves accuracy by 98.11\%, 26.74\%, and 64.17\%,
respectively.

Figure~\ref{fig:box_plot_accuracy} presents the accuracy box plot across the hyperparameter search space for these applications (higher is better).  
The central rectangle covers the first-third quartile
\texttt{$(Q3 - Q1)$} and the whiskers span from \texttt{$(Q3 + 2 * IQR)$} to \texttt{$(Q1 - 2 * IQR)$}.
The red notch indicates the accuracy metric from the model trained using the default
settings and colored diamonds show outliers beyond the whiskers.
The blue triangle at the top identifies the accuracy percentage reported by Seneca.

The model output quality results across applications, show that prediction
accuracy (for a given dataset) is dramatically
affected by hyperparameter settings.  Predictably, the default settings are
near the ``good'' end of the spectrum, however, Seneca is able to find the
parameterization that improves output quality over the default settings
in each case.

To investigate the potential impact of Seneca's 80/20 percent data split for the classification
applications, we next evaluate the quality of the output generated from 
each when we consider different 80/20 random splits.
To enable this, we run Seneca 30 times to obtain execution time, memory use, 
and best accuracy score. We report the
mean and standard deviation (in parentheses) 
for execution time and memory use across runs,
and the best accuracy score in
Table~\ref{tab:exec_memory}. Each pair of rows shows the results for two different random splits.
Our earlier results use input 1; this table adds results for a second, 80/20 
random split of the input (we also considered other random splits, which we omit for brevity, 
and the results are similar).  The performance and Seneca score is similar across splits. 
This result indicates that for these applications, 
users can repeatedly employ the recommended models for inference on other datasets or splits,
to amortize the cost of using Seneca.
